### 2023-05-29 paper/Chapters/Chapter1.tex
Final strokes

As reported by \textcite{duisterhofTartanCalibIterativeWideAngle2022} on
\citedate{duisterhofTartanCalibIterativeWideAngle2022}, the current state-of-the-art
methods
\citep{olsonAprilTagRobustFlexible2011, schopsWhyHaving102020,
	krogiusFlexibleLayoutsFiducial2019}
fail on images with high distortion.
\cite{duisterhofTartanCalibIterativeWideAngle2022} suggested an iterative
the approach of image undistortion and target reprojection, achieving the superior
robustness to the noise than the state-of-the-art methods because the feature
detection is performed on the undistorted image.
% \todo{Probably remove the emphasis on TartanCalib, if I am not going to compare
% with it}

@@

The objective of this research is to improve the accuracy of the camera
calibration by finding previously undetected features on the calibration
board \todo{Probably rephrase if I won't compare the actual calibration}. For
that, we formulate the set of research questions:
\begin{itemize}
	\item How to find additional features on the calibration board which were not
	      detected by the feature detector?
	\item How to filter out falsely detected features?
	      \todo{Add more}
\end{itemize}

@@

This paper has the following structure: in \cref{related_work}, we will describe
various subtopics of the camera calibration, and outline the current state-of-the-art solutions.
In \cref{cha:background}, we will describe the key concepts of the camera
calibration. Also, there we provide additional details on the underlying math,
for example, the derivation of the division model invertion. In
\cref{cha:approach}, we will describe the steps of the proposed method in
details, including the feature detection, initial camera calibration, camera
parameters estimation via minimization of the reprojection error, and the
process of searching and filtering the previously undetected features. In
\cref{Experiments}, we will describe the metrics we used, provide details about
the datasets, show the results of each of the algorithm's steps, and evaluate it
on multiple metrics. Finally, in \cref{cha:conclusions}, we will summarize the
results and outline future work.

>>>>

As reported by \textcite{duisterhofTartanCalibIterativeWideAngle2022} on
\citedate{duisterhofTartanCalibIterativeWideAngle2022}, the current state-of-the-art
methods
\citep{olsonAprilTagRobustFlexible2011, schopsWhyHaving102020,
	krogiusFlexibleLayoutsFiducial2019}
fail on images with high distortion.
\cite{duisterhofTartanCalibIterativeWideAngle2022} suggested an iterative
the approach of image undistortion and target reprojection, achieving the superior
robustness to the noise than the state-of-the-art methods because the feature
detection is performed on the undistorted image.

@@

The objective of this research is to add additional constrains to the camera
calibration by finding previously undetected features on the calibration
board. For that, we formulate the set of research questions:
\begin{itemize}
	\item How to find additional features on the calibration board which were not
	      detected by the feature detector?
	\item How to filter out falsely detected features?
  \item Is there a need for finding additional features on the calibration
    board? Are all of the points detected?
\end{itemize}

@@

This paper has the following structure: in \cref{related_work}, we will describe
various subtopics of the camera calibration, and outline the current state-of-the-art solutions.
In \cref{cha:background}, we will describe the key concepts of camera
calibration. Also, there we provide additional details on the underlying math,
for example, the derivation of the division model inversion. In
\cref{cha:approach}, we will describe the steps of the proposed method in
detail, including feature detection, initial camera calibration, camera
parameters estimation via minimization of the reprojection error, and the
process of searching and filtering the previously undetected features. In
\cref{Experiments}, we will describe the metrics we used, provide details about
the datasets, show the results of each of the algorithm's steps, and evaluate it
on multiple metrics. Finally, in \cref{cha:conclusions}, we will summarize the
results and outline future work.

### 2023-05-29 paper/Chapters/Chapter4.tex
WIP
Final strokes

For feature detection, we used the approach, proposed by
\cite{geigerAutomaticCameraRange2012} (see \cref{sec:boards_features_detection,
	sec:feature_detection}).

@@

In this work, we obtained the camera calibration in two steps: first, we
used the approach, proposed by \cite{scaramuzzaToolboxEasilyCalibrating2006} to
obtain the initial approximation of the camera parameters. Then, we used 
optimization to minimize the reprojection error between the board and the
back-projected corners. A division model is very powerful and can
express a wide range of distortions
\citep{prittsMinimalSolversRectifying2021}. In this work, we used a model with
two parameters to allow for better flexibility during the optimization.

@@

\begin{equation}
	R = R(\theta_x) R(\theta_y) R(\theta_z)
\end{equation} \todo{Is the notation ok?}

@@

For \(K\) we used the placeholder values. \(f\) was set to a constant value for
the typical consumer camera, and \(c_x, c_y\) were set to the center of the
image.
The correct values will be computed during the optimization \cref{sub:optimization}.

@@

Only \cref{eq:scaramuzza_system_3} is linear in the unknowns. Each point gives a
single equation. Now, by rewriting the equation in the matrix form
\(M \cdot H = 0\), where \(H = \begin{pmatrix}
	r_{11}, r_{12}, r_{12}, r_{22}, t_1, t_2
\end{pmatrix}\) we get:

\begin{equation}
	M = \begin{bmatrix}
		-\widehat{v}_1 X_1 & -\widehat{v}_1 Y_1 & -\widehat{u}_1 X_1 & -\widehat{u}_1 Y_1 & -\widehat{v}_1 & -\widehat{u}_1 \\
		\vdots             & \vdots             & \vdots             & \vdots             & \vdots         & \vdots         \\
		-\widehat{v}_N X_N & -\widehat{v}_N Y_N & -\widehat{u}_N X_N & -\widehat{u}_N Y_N & -\widehat{v}_N & -\widehat{u}_N
	\end{bmatrix}
\end{equation}, where \(N\) is the number of points.

@@

Now, to solve for \(t_1\) and \(t_2 \), first find possible values for
\(t_2 ^{2}\):
\begin{align}
	- \frac{r_{11} r_{12} + r_{12} r_{22}}{t_2 }                  & = t_1 &  &
	\text{Solve eq. (\ref{eq:orthonormality_1}) for \(t_1\)}
	\label{eq:orthonormality_4}                                                \\
	\\
	\sqrt{\lambda^{2} \left( \frac{r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2}\right)}                               & = 1   &  &
	\text{Substitute into eq. (\ref{eq:orthonormality_2})}                     \\
	\lambda^{2} (\frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2})                                      & = 1   &  &
	\text{Square both sides}                                                   \\
	\lambda^{2} (\frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2} - r_{12}^{2} - r_{22}^{2} - t_2 ^{2}) & = 0   &  &
	\text{Subtract eq. (\ref{eq:orthonormality_3})}
	\label{eq:orthonormality_5}
	\\
	\frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2} - r_{12}^{2} - r_{22}^{2} - t_2 ^{2}  & = 0   &  &
	\text{Divide both by \(\lambda ^{2}\)}                                     \\
	t_2 ^{4} - (r_{11}^{2} + r_{12}^{2} - r_{12}^{2} -
	r_{22}^{2}) t_2 ^{2} - (r_{11} r_{12} + r_{12} r_{22})^{2}    & = 0   &  &
	\text{Multiply by \(t_2 ^{2}\)}
	\label{eq:orthonormality_6}
\end{align}

@@

This issue also occurred with random small initial values, due to the best
approximation for \(R\) which minimizes \(L\) when the distance from the
back-projected board from measure one was high (i.e., the \(t\) was far from
the true value) being the degenerate solution\todo{Am I using the term
	correctly?}. To avoid this issue, we first optimized only the \(t\),
until the loss function converged, meaning that \(t\) is close to the true
value.

@@

Often, not all of the board's corners were detected initially. Firstly, we
assumed that the whole board was detected, and imputed the missing points in the
board space \cref{fig:extended_board_img}. Then, we tried extending the board
points.

@@

	\begin{subfigure}{0.45\textwidth}
		\centering
		\begin{minipage}{\textwidth}
			\includegraphics[width=0.24\textwidth]{./template_1_a.png}
			\includegraphics[width=0.24\textwidth]{./template_1_b.png}
			\includegraphics[width=0.24\textwidth]{./template_1_c.png}
			\includegraphics[width=0.24\textwidth]{./template_1_d.png}
		\end{minipage}
		\caption{Corner prototype 1 (A, B, C, D)}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.45\textwidth}
		\centering
		\begin{minipage}{\textwidth}
			\includegraphics[width=0.24\textwidth]{./template_2_a.png}
			\includegraphics[width=0.24\textwidth]{./template_2_b.png}
			\includegraphics[width=0.24\textwidth]{./template_2_c.png}
			\includegraphics[width=0.24\textwidth]{./template_2_d.png}
		\end{minipage}
		\caption{Corner prototype 2 (A, B, C, D)}
	\end{subfigure}

>>>>

For feature detection, we used the approach, proposed
by~\cite{geigerAutomaticCameraRange2012} (see \cref{sec:boards_features_detection,
	sec:classifier}).

@@

In this work, we obtained the camera calibration in two steps: first, we
used the approach, proposed by~\cite{scaramuzzaToolboxEasilyCalibrating2006} to
obtain the initial approximation of the camera parameters. Then, we used
optimization to minimize the reprojection error between the board and the
back-projected corners. A division model is very powerful and can
express a wide range of distortions
\citep{prittsMinimalSolversRectifying2021}. In this work, we used a model with
two parameters to allow for better flexibility during the optimization.

@@

\begin{equation}
	R = R(\theta_x) R(\theta_y) R(\theta_z)
\end{equation}

@@

For \(K\) we used the placeholder values.\ \(f\) was set to a constant value for
the typical consumer camera, and \(c_x, c_y\) were set to the center of the
image.
The correct values will be computed during the optimization \cref{sub:optimization}.

@@

Only \cref{eq:scaramuzza_system_3} is linear in the unknowns. Each point gives a
single equation. Now, by rewriting the equation in the matrix form
\(M \cdot H = 0\), where
\[
	H = \begin{pmatrix}
		r_{11}, r_{12}, r_{12}, r_{22}, t_1, t_2
	\end{pmatrix}
\]
we get:
\begin{equation}
	M = \begin{bmatrix}
		-\widehat{v}_1 X_1 & -\widehat{v}_1 Y_1 & -\widehat{u}_1 X_1 & -\widehat{u}_1 Y_1 & -\widehat{v}_1 & -\widehat{u}_1 \\
		\vdots             & \vdots             & \vdots             & \vdots             & \vdots         & \vdots         \\
		-\widehat{v}_N X_N & -\widehat{v}_N Y_N & -\widehat{u}_N X_N & -\widehat{u}_N Y_N & -\widehat{v}_N & -\widehat{u}_N
	\end{bmatrix}
\end{equation}, where \(N\) is the number of points.

@@

Now, to solve for \(t_1\) and \(t_2 \), first find possible values for
\(t_2 ^{2}\):
\begin{align}
	- \frac{r_{11} r_{12} + r_{12} r_{22}}{t_2 }                  & = t_1 &  &
	\text{Solve eq. (\ref{eq:orthonormality_1}) for \(t_1\)}
	\label{eq:orthonormality_4}                                                \\
	\\
  \sqrt{\lambda^{2} \left( \frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2}\right)}                               & = 1   &  &
	\text{Substitute into eq. (\ref{eq:orthonormality_2})}                     \\
  \lambda^{2} \left( \frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}}\right) +
	r_{11}^{2} + r_{12}^{2}                                      & = 1   &  &
	\text{Square both sides}                                                   \\
  \lambda^{2} \left(\frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} \right) +
	r_{11}^{2} + r_{12}^{2} - r_{12}^{2} - r_{22}^{2} - t_2 ^{2} & = 0   &  &
	\text{Subtract eq. (\ref{eq:orthonormality_3})}
	\label{eq:orthonormality_5}
	\\
	\frac{(r_{11} r_{12} + r_{12} r_{22})^{2}}{t_2 ^{2}} +
	r_{11}^{2} + r_{12}^{2} - r_{12}^{2} - r_{22}^{2} - t_2 ^{2}  & = 0   &  &
	\text{Divide both by \(\lambda ^{2}\)}                                     \\
	t_2 ^{4} - (r_{11}^{2} + r_{12}^{2} - r_{12}^{2} -
	r_{22}^{2}) t_2 ^{2} - (r_{11} r_{12} + r_{12} r_{22})^{2}    & = 0   &  &
	\text{Multiply by \(t_2 ^{2}\)}
	\label{eq:orthonormality_6}
\end{align}

@@

This issue also occurred with random small initial values, due to the best
approximation for \(R\) which minimizes \(L\) when the distance from the
back-projected board from measure one was high (i.e., the \(t\) was far from
the true value) being the degenerate solution. To avoid this issue, we first optimized only the \(t\),
until the loss function converged, meaning that \(t\) is close to the true
value.

@@

Often, not all of the board's corners were detected initially. Firstly, we
assumed that the whole board was detected, and imputed the missing points in the
board space (\cref{fig:extended_board_img}). Then, we tried extending the board
points.

@@

	\begin{subfigure}{0.45\textwidth}
		\centering
		\begin{minipage}{\textwidth}
			\includegraphics[width=0.24\textwidth]{./template_1_a.png}
			\includegraphics[width=0.24\textwidth]{./template_1_b.png}
			\includegraphics[width=0.24\textwidth]{./template_1_c.png}
			\includegraphics[width=0.24\textwidth]{./template_1_d.png}
		\end{minipage}
		\caption{Corner prototype 1 (A, B, C, D)}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.45\textwidth}
		\centering
		\begin{minipage}{\textwidth}
			\includegraphics[width=0.24\textwidth]{./template_2_a.png}
			\includegraphics[width=0.24\textwidth]{./template_2_b.png}
			\includegraphics[width=0.24\textwidth]{./template_2_c.png}
			\includegraphics[width=0.24\textwidth]{./template_2_d.png}
		\end{minipage}
		\caption{Corner prototype 2 (A, B, C, D)}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\includegraphics[width=\textwidth]{10.png}
		\caption{Input image}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\includegraphics[width=\textwidth]{11.png}
		\caption{Corner likelihood}
	\end{subfigure}

### 2023-05-29 paper/Chapters/Chapter5.tex
Final strokes

\textbf{Kalibr} \citep{mayeSelfsupervisedCalibrationRobotic2013} contains several established datasets that are commonly used for testing
the accuracy of camera calibration frameworks: Double Sphere
\cite{usenkoDoubleSphereCamera2018}, EuRoC \cite{burriEuRoCMicroAerial2016}, TUM
VI \cite{schubertTUMVIBenchmark2018}, and ENTANIYA 1
\cite{Calibration250degFisheye}.
The Kalibr calibration framework was used in the
original publications cited above, hence the name of the dataset.
As a calibration pattern, AprilGrid with 6x6 tags of 88 mm size was used.
In total, the datasets contain approximately 800 images.

@@

\textbf{UZH} \citep{AreWeReady} is a dataset of approximately 800 images
collected using the following cameras:
As a calibration pattern, AprilGrid with 4x5 tags of 75 mm size was used.
The dataset contains approximately 800 images.

\textbf{OV} \citep{lochmanBabelCalibUniversalApproach2021} is a dataset of
approximately 1400 images. It was collected using eight stereo cameras.
As a calibration pattern, the checkerboard pattern with 9x6 tags of 22 mm size
was used.

@@

\begin{minipage}{0.5\linewidth}
	To get accurate predictions for the possibly missing points, we need
	very accurate camera calibration. We used the reprojection error as a metric
	for the calibration quality.
	To calculate the reprojection error, we have to perform the
	rootfinding \cref{subsub:back_projection_using_the_division_model}. We assume
	that all of the points lay within the image, or they get out of the borders
	a little bit (we use the constant \(1.1\) of the maximum radius). If the
	camera calibration is poor, the root-finding might fail. On the initial
	calibration the number of the failed root-findings was much higher.
\end{minipage}
\hfill
\begin{minipage}{0.4\linewidth}
	\resizebox{\linewidth}{!}{
		\begin{tikzpicture}
			\pie[rotate=90, explode=0.1, sum=auto, color={green!60, blue!60, red!60}, text=legend]{
				115/Correctly detected,
				80/Reprojection error > 10,
				254/Not correctly detected
			}
		\end{tikzpicture}
	}
	\captionsetup{width=.9\linewidth}
	\captionof{figure}{Initial calibration}
	\resizebox{\linewidth}{!}{
		\begin{tikzpicture}
			\pie[rotate=90, explode=0.1, sum=auto, color={green!60, blue!60, red!60}, text=legend]{
				658/Correctly detected,
				19/Reprojection error > 10,
				115/Not correctly detected
			}
		\end{tikzpicture}
	}
	\captionsetup{width=.9\linewidth}
	\captionof{figure}{Final calibration}
\end{minipage}

@@

We removed 20\% of the points from the original board, and then tried to recover
them.

@@

Occlusions pose additional complications for the feature detection. We added an
image on top of the board, and tried to recover additional points. Often the
initial feature detector failed to detect points near the occlusion.

@@

\begin{minipage}[t]{0.3\linewidth}
	Lastly, we recovered the points that were not detected by the initial feature
	detector.
	From a visual inspection, most of the points were recovered correctly
	\cref{fig:recovered_good_points}.
	On the top image, for example, there are green points close to the edge.
	Those corners weren't detected initially due to high distortion. On the bottom
	image, the algorithm found another row of the corners, which weren't detected previously.
\end{minipage}
\hfill
\begin{minipage}[t]{0.6\linewidth}
	\includegraphics[width=\linewidth]{refined_corners1.png}
	\includegraphics[width=\linewidth]{refined_corners2.png}

>>>>

\textbf{Kalibr} \citep{mayeSelfsupervisedCalibrationRobotic2013} contains several established datasets that are commonly used for testing
the accuracy of camera calibration frameworks: Double Sphere
\cite{usenkoDoubleSphereCamera2018}, EuRoC \cite{burriEuRoCMicroAerial2016}, TUM
VI \cite{schubertTUMVIBenchmark2018}, and ENTANIYA 1
\cite{Calibration250degFisheye}.
The Kalibr calibration framework was used in the
original publications cited above, hence the name of the dataset.
As a calibration pattern, AprilGrid with  \(6\times_6\) tags of 88 mm size was used.
In total, the datasets contain approximately 800 images.

@@

\textbf{UZH} \citep{AreWeReady} is a dataset of approximately 800 images
collected using the following cameras:
As a calibration pattern, AprilGrid with \(4\times_5\) tags of 75 mm size was used.
The dataset contains approximately 800 images.

\textbf{OV} \citep{lochmanBabelCalibUniversalApproach2021} is a dataset of
approximately 1400 images. It was collected using eight stereo cameras.
As a calibration pattern, the checkerboard pattern with \(9\times_6\) tags of 22 mm size
was used.

@@

\begin{minipage}{0.5\linewidth}
	To get accurate predictions for the possibly missing points, we need
	very accurate camera calibration. We used the reprojection error as a metric
	for the calibration quality.
	To calculate the reprojection error, we have to perform the
	rootfinding \cref{subsub:back_projection_using_the_division_model}. We assume
	that all of the points lay within the image, or they get out of the borders
	a little bit (we use the constant \(1.1\) of the maximum radius). If the
	camera calibration is poor, the root-finding might fail. On the initial
	calibration the number of failed root-findings was much higher.
\end{minipage}
\hfill
\begin{minipage}{0.4\linewidth}
	\resizebox{\linewidth}{!}{
		\begin{tikzpicture}
			\pie[rotate=90, explode=0.1, sum=auto, color={green!60, blue!60, red!60}, text=legend]{
				115/Correctly detected,
				80/Reprojection error > 10,
				254/Not correctly detected
			}
		\end{tikzpicture}
	}
	\captionsetup{width=.9\linewidth}
	\captionof{figure}{Initial calibration}
	\resizebox{\linewidth}{!}{
		\begin{tikzpicture}
			\pie[rotate=90, explode=0.1, sum=auto, color={green!60, blue!60, red!60}, text=legend]{
				658/Correctly detected,
				19/Reprojection error > 10,
				115/Not correctly detected
			}
		\end{tikzpicture}
	}
	\captionsetup{width=.9\linewidth}
	\captionof{figure}{Final calibration}
\end{minipage}

@@

We removed 20\% of the points from the original board and then tried to recover
them.

@@

Occlusions pose additional complications for feature detection. We added an
image on top of the board and tried to recover additional points. Often the
initial feature detector failed to detect points near the occlusion.

@@

\begin{minipage}[t]{0.3\linewidth}
	Lastly, we recovered the points that were not detected by the initial feature
	detector.
	From a visual inspection, most of the points were recovered correctly
	\cref{fig:recovered_good_points}.
	On the top image, for example, there are green points close to the edge.
	Those corners weren't detected initially due to high distortion. On the bottom
	image, the algorithm found another row of corners, which weren't detected previously.
\end{minipage}
\hfill
\begin{minipage}[t]{0.6\linewidth}
	\includegraphics[width=\linewidth]{refined_corners1.png}
	\includegraphics[width=\linewidth]{refined_corners2.png}

### 2023-05-29 paper/paper.tex
WIP
Final strokes

@@

\begin{acknowledgements}
	\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
  I wish to thank the Armed Forces of Ukraine for protecting us. Thanks to those
  brave people I was able to study and work on this thesis. I want to thank my
	% The acknowledgments and the people to thank go here, don't forget to include your project advisor\ldots

@@

% \tableofcontents % Prints the main table of contents
\todo{Enable}
%
% \listoffigures % Prints the list of figures
\todo{Enable}
%
% \listoftables % Prints the list of tables
\todo{Enable}

@@

% \dedicatory{For/Dedicated to/To my\ldots}
\todo{Enable}

>>>>

  The code for this paper is available at
  \url{https://github.com/anstadnik/camera_calibration}.

@@

\begin{acknowledgements}
	\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
	I wish to thank the Armed Forces of Ukraine for protecting us. Thanks to those
	brave people I was able to study and work on this thesis. I want to thank my
	supervisor, Dr. James Pritts, for his guidance and support throughout the whole
	process of writing this thesis. I also want to express my gratitude to all of
	the people who encouraged me to keep going and inspired me to do my best: my
	groupmates, my friends, and my girlfriend.
	% The acknowledgments and the people to thank go here, don't forget to include your project advisor\ldots

@@

\tableofcontents % Prints the main table of contents
%
% \listoffigures % Prints the list of figures
%
% \listoftables % Prints the list of tables

@@

% \dedicatory{For/Dedicated to/To my\ldots}
